The ChEMBL database as Linked Open Data

### 2012-04-05 article.tex
add address for my university
Completed a draft introduction into the graph structure

\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands}
\address[B]{Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden}
\address[C]{University of Queensland}
\address[D]{Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A.}
\address[E]{ChEBI, European Bioinformatics Institute}
\address[F]{ChEMBL, European Bioinformatics Institute}
\address[G]{Indiana University}

@@

@@

Compounds ... Three types of compounds, peptides, small chemicals, ...

\begin{footnotesize}
\begin{verbatim}
m41:inchikey
 a cheminf:CHEMINF_000059 ;
 cheminf:SIO_000300
   "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .

@@

chemblid:CHEMBL406142
 owl:equivalentClass mol:m41 .

mol:m41
 rdfs:label "CHEMBL406142" , 
   "Bis(3-[1 .... yl]-propionamide)" ;
 rdfs:subClassOf cheminf:CHEMINF_000000 ;
 cheminf:CHEMINF_000200
  m41:inchikey , m41:smiles , m41:inchi ;
 owl:equivalentClass 
  <http://rdf.openm .... 66+,67+,68+/m1/s1> ,
  <http://bio2rdf.org/chebi:100041> ,
  chemblid:CHEMBL406142 .
\end{verbatim}
\end{footnotesize}

The molecular properties available from ChEMBL are now too exposed in triple format,
and like the InChI and InChIKey, these are provided using the CHEMINF ontology structure:

@@

FIXME: add ChemSpider IDs with CHEMINF.

For documents little information is replicated from the database, taking advantage
of PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs). URIs for the
latter can be resolved online, providing curated information on the identified
documents. For PubMed articles ...

@@

res:r1
 a bibo:Article ;
 rdfs:seeAlso <http://bio2rdf.org/pubmed:14695813> ;
 dc:date "2004" ;
 dc:isPartOf
  journal:j6c706049c2e08871b7c46a6528065736 ;
 bibo:issue "1" ;
 bibo:pageEnd "9" ;
 bibo:pageStart "1" ;
 bibo:pmid "14695813" ;
 bibo:volume "47" .
\end{verbatim}
\end{footnotesize}

>>>>

\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, ADDRESS, Maastricht,\\ The Netherlands}
\address[B]{Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden}
\address[C]{University of Queensland, St Lucia, Qld 4072, Australia}
\address[D]{Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A.}
\address[E]{ChEBI, European Bioinformatics Institute}
\address[F]{ChEMBL, European Bioinformatics Institute}
\address[G]{Indiana University}

@@

More than five thousand different activity types are captured by the ChEMBL database.
The top five types are Potency (43\%), IC50 (13\%), MIC (4.6\%), Inhibition (3.7\%),
and Ki (3.6\%). The activity types in ChEMBL-RDF are currently not available as, or
using, an ontology.

@@

For drug discovery, the drugs themselves are the main topic of study.
ChEMBL contains many different drug types, mostly small molecules,
but also peptides, proteins, oligosaccharides, oligonucleotides, and
even antibodies. All compounds are represented as classes, following
the design of the CHEMINF ontology. The compound typing is expressed
by subclassing other classes. For example, proteins subclass the
Protein concept from the PRotein Ontology (PR\_000000001), small
molecules subclass the Chemical Entity concept from the CHEMINF
ontology (CHEMINF\_000000), and oligosaccharides and oligonucleotides
subclass their respective matches in the CHEBI ontology (CHEBI\_50699
and CHEBI\_7754). For each drug compound the name and synonyms are
provided as labels. When InChI and InChIKeys are available for a drug, then these are
provided via the CHEMINF formalism (here abbreviated):

\begin{footnotesize}
\begin{verbatim}
mol:m41
 rdfs:label "CHEMBL406142" , 
   "Bis(3-[1 .... yl]-propionamide)" ;
 rdfs:subClassOf cheminf:CHEMINF_000000 ;
 cheminf:CHEMINF_000200
  m41:inchikey , m41:smiles , m41:inchi ;

m41:inchikey
 a cheminf:CHEMINF_000059 ;
 cheminf:SIO_000300
   "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .

@@

chemblid:CHEMBL406142
 owl:equivalentClass mol:m41 .
\end{verbatim}
\end{footnotesize}

For small molecules, molecular properties are often available from ChEMBL, and as of ChEMBL-RDF 13
these too are exposed in triple format. Like the InChI and InChIKey, these are provided using the
CHEMINF ontology approach. Here is, for example, the ALogP value:

@@

For documents little information is replicated from the database, taking advantage
of PubMed Identifiers (PMIDs) and Digital Object Identifiers (DOIs). URIs for the
latter can be resolved online, providing curated information on the identified
documents. For each paper, basic properties are provided using the BIBO
ontology:

@@

res:r1
 a bibo:Article ;
 dc:date "2004" ;
 dc:isPartOf
  journal:j6c706049c2e08871b7c46a6528065736 ;
 bibo:issue "1" ;
 bibo:pageEnd "9" ;
 bibo:pageStart "1" ;
 bibo:pmid "14695813" ;
 bibo:volume "47" .
\end{verbatim}
\end{footnotesize}
